---
title: "WRU_squadscraping"
author: "BY"
date: "`r Sys.Date()`"
output: html_document
---
Install and load the libraries we need. Rvest is for scraping, which I use with the Polite package.

```{r}
#install.packages("rvest")
#install.packages("dplyr")
#install.packages("polite")
#install.packages("gh")
#install.packages("usethis")
# Install devtools if not already installed
install.packages("devtools")
install.packages('polite', dependencies = TRUE, repos='http://cran.rstudio.com/')

library(rvest)
library(polite)
library(dplyr)
library(tidyr)
```
Helpful pages on these packages: https://personales.upv.es/algarsal/R-tutorials/07_Tutorial-7_R-dplyr-tidyr.html
https://dmi3kno.github.io/polite/

Define the url and extract the html
```{r}
host <- "https://en.wikipedia.org/wiki/2008_Six_Nations_Championship_squads"
session <- polite::bow(host, force = TRUE)

url <- "https://en.wikipedia.org/wiki/2008_Six_Nations_Championship_squads"
session <- polite::bow(url, force = TRUE)

webpage <- read_html(url)
```

First, I need to inspect the web page structure to locate the table for the Wales squad. Wikipedia tables often have class names or other identifiers. We’ll look for a table under the Wales section.

Extract all tables from the webpage.
```{r}
tables <- html_nodes(webpage, "table")
```

Pull out the Wales table.
```{r}
WRUSquad_2008 <- html_table(tables[[6]], fill = TRUE)
```

Convert the list to a data frame.
```{r}
df_WRUSquad_2008 <- as.data.frame(WRUSquad_2008)
```

Clean up the column names and data as necessary.
```{r}
colnames(df_WRUSquad_2008) <- c("Player", "Position", "DOB", "Caps", "Club")
```

Remove empty rows.
```{r}
df_WRUSquad_2008 <- df_WRUSquad_2008 %>% filter(Player != "") 
```

Split out DOB into 3 columns
```{r}
library(stringr)
df_WRUSquad_2008 <- df_WRUSquad_2008 %>% separate(DOB, c('DOB_Day', 'DOB_Month', 'DOB_Year'))

#df_WRUSquad_2008 <- as.data.frame(lapply(df_WRUSquad_2008$DOB_Year))
 
print(df_WRUSquad_2008)
```

I want to add a Squad year column to all rows
```{r}
C_len <- nrow(df_WRUSquad_2008)
C_len

Squad_Year <-rep(c(2008),times=C_len, as.numeric)
df1<-data.frame(Squad_Year)
Squad_Year

df_WRUSquad_2008 = data.frame(df_WRUSquad_2008, Squad_Year)
print(df_WRUSquad_2008)
```


Replace all spaces in the Player column with underscores, using the tidyverse package, and remove the full stops. The full stop is a special character in R, so we need to escape it using [].
```{r}
df_WRUSquad_2008 <- df_WRUSquad_2008 %>%
   mutate(Player = str_replace_all(Player, "[.]", ""))

df_WRUSquad_2008 <- df_WRUSquad_2008 %>%
    mutate(Player = str_replace_all(Player, " ", "_"))
```

Calculate the age of players in the Squad_Year:
Check dataframe dimensions
```{r}
print(nrow(df_WRUSquad_2008))  # Should print 30
```

```{r}
#check missing values
df_WRUSquad_2008 <- df_WRUSquad_2008 %>%
  filter(!is.na(Squad_Year) & !is.na(DOB_Year))
```

```{r}
# Verify column names
print(colnames(df_WRUSquad_2008))
```

```{r}
# Check data types
print(str(df_WRUSquad_2008))
```

```{r}
# Convert DOB_Year to numeric
df_WRUSquad_2008$DOB_Year <- as.numeric(as.character(df_WRUSquad_2008$DOB_Year))
```

```{r}
#Create Age Column and calculate player age at time of Six Nations tournament
df_WRUSquad_2008 <- df_WRUSquad_2008 %>%
  mutate(Age = Squad_Year - DOB_Year)

#rename column
#df_WRUSquad_2008 <- df_WRUSquad_2008 %>%
#  rename(Age = age)
```

Display and write cleaned table
```{r}
print(df_WRUSquad_2008)
write.csv(df_WRUSquad_2008, file = "df_WRUSquad_2008.csv")
```
Repeat for other years

2009

```{r}
url <- "https://en.wikipedia.org/wiki/2009_Six_Nations_Championship_squads"

webpage <- read_html(url)
tables <- html_nodes(webpage, "table")

```

I want to generate random PlayerID codes
